---
title: "R Notebook"
output: html_notebook
---

The [R plugin](https://www.jetbrains.com/help/pycharm/r-plugin-support.html) for IntelliJ-based IDEs provides
handy capabilities to work with the [R Markdown](https://www.jetbrains.com/help/pycharm/r-markdown.html) files.
To [add](https://www.jetbrains.com/help/pycharm/r-markdown.html#add-code-chunk) a new R chunk,
position the caret at any line or the code chunk, then click "+".

The code chunk appears:
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)
library(tidyverse)
library(Metrics)
library(astsa)
library(tseries)
library(readxl)
library(tsutils)
library(forecast)
library(urca)
library(lmtest)
library(vars)
library(tsDyn)
```

Type any R code in the chunk, for example:
```{r}
# import data + plot series:
HSG_train <- read_excel("C:\\Users\\ADMIN\\RProjects\\Viet_Nam_Steel_stock_analysis-Time-Series-\\Stock Price - HSG - train.xlsx")
HPG_train <- read_excel("C:\\Users\\ADMIN\\RProjects\\Viet_Nam_Steel_stock_analysis-Time-Series-\\Stock Price - HPG - train.xlsx")
NKG_train <- read_excel("C:\\Users\\ADMIN\\RProjects\\Viet_Nam_Steel_stock_analysis-Time-Series-\\Stock Price - NKG - train.xlsx")
```
```{r}
HSG <- HSG_train$`<Close>`
HPG <- HPG_train$`<Close>`
NKG <- NKG_train$`<Close>`
```
```{r}
summary(ca.jo(data.frame(HSG, HPG, NKG), type = 'trace'))
```
```{r}
summary(ca.jo(data.frame(HSG, HPG, NKG), type = 'eigen'))
```
# lambda_trace
          test 10pct  5pct  1pct
r <= 1 | 15.83 15.66 17.95 23.52
r = 0  | 36.18 28.71 31.52 37.22

r = 0:  test = 36.18 > 31.52  -> reject H0: Number of cointegrate > 0
r <= 1: test = 15.83 < 17.95  -> Not reject H0: Number of cointegrate =< 1 (VECM ở dưới)

# lambda_max_eigen
          test 10pct  5pct  1pct
r = 0  | 20.35 18.90 21.07 25.75

r = 0:  test = 20.35 > 21.07  -> Not reject H0: Number of cointegrate = 0


```{r}
summary(lm(HPG ~ HSG + NKG))
```
```{r}
resid.HPG <- resid(summary(lm(HPG ~ HSG + NKG)))
summary(ur.df(resid.HPG, type = 'none'))
```
```{r}
summary(lm(diff(HPG) ~ diff(HSG) + diff(NKG) + resid.HPG[1:(length(resid.HPG) - 1)]))
```
# VAR
```{r}
HPG.d <- diff(HPG)
HSG.d <- diff(HSG)
NKG.d <- diff(NKG)
data_diff <- data.frame(HPG.d, HSG.d, NKG.d)
data_non_diff <- data.frame(HPG, HSG, NKG)
```

```{r}
# grangertest(HPG.d ~ HSG.d, order = 1 )
# grangertest(HSG.d ~ HPG.d, order = 1 )
# grangertest(HPG.d ~ NKG.d, order = 1 )
# grangertest(NKG.d ~ HPG.d, order = 1 )
# grangertest(HSG.d ~ NKG.d, order = 1 )
# grangertest(NKG.d ~ HSG.d, order = 1 )
```
```{r}
var_diff_lag1 <- VAR(data_diff, p= 1, type = 'const')
summary(var_diff_lag1)
```
# Test for model
Significant coefficients

No serrial correlation:         H0 :No serrial correlation
                                H1 :Serrial correlation
```{r}
serial.test(var_diff_lag1)
```
# p - value = 0.009 -> Reject H0 -> Serial Correlation
```{r}
plot(predict(var_diff_lag1))
```
# IRF
```{r}
irf(var_diff_lag1)
```
```{r}
plot(irf(var_diff_lag1))
```
# FEDV
```{r}
plot(fevd(var_diff_lag1))
```
```{r}
VARselect(data_diff, lag.max = 4, type = 'const')
```
```{r}
VARselect(data_diff, lag.max = 4, type = 'none')
```
```{r}
VECM(data.frame(HSG, HPG, NKG), lag= 1, include = 'none')
```
```{r}
VECM(data.frame(HPG, HSG, NKG), lag= 1, include = 'const')
```
# VAR model in difference using tsDyn
```{r}
VAR_diff_lag1 = lineVar(data = data_diff, lag = 1, include = 'const', model = 'VAR', I = 'diff')
VAR_diff_lag1
```
```{r}
m.lev <- as.matrix(data_non_diff)
nr_lev <- nrow(data_non_diff)

# tạo diff
df.diff <- diff(as.matrix(data_non_diff), lag = 1)
colnames(df.diff) <- c('d.HPG', 'd.HSG', 'd.NKG')
m.diff <- as.matrix(df.diff)

# estimation
vare_diff <- VAR(df.diff, p = 1, type = 'const')

# forecast of differenced data
varf_diff <- predict(vare_diff, n.ahead = 10)
plot(varf_diff)
fanchart(varf_diff)
```
```{r}
# recover lev forecast
m.varf_lev_ft <- rbind(m.lev, matrix(NA, 10, 3))
m.ft_df <- do.call(cbind,lapply(varf_diff$fcst, function(x) x[,'fcst']))

# growth to level
for(h in (nr_lev+1):(nr_lev + 10)) {
    hf <- h - nr_lev
    m.varf_lev_ft[h,] <- m.varf_lev_ft[h-1,] + m.ft_df[hf,]
}

# Draw Graph

for(i in 1:3) {
    df <- ts(m.varf_lev_ft[1:499,i], start = 2021, frequency = 250)
    forecast_df <- ts(m.varf_lev_ft[500:509,i], start = 2023, frequency = 250)

    plot(df, main="VAR model ~ time")
    lines(forecast_df, col = c('blue'))
    legend("topleft", legend=c("Actual", "Forecast 2023"),
       col=c("black", "blue"), lty=1:2, cex=0.8)
}
```
```{r}
ts(m.varf_lev_ft[1:499,1], start = c(1, 1), end = c(2, 249))
```
```{r}
var_diff_lag1$varresult
```
```{r}
# HPG
accuracy(var_diff_lag1$varresult[[1]])
```
```{r}
# HSG
accuracy(var_diff_lag1$varresult[[2]])
```
```{r}
# NKG
accuracy(var_diff_lag1$varresult[[3]])
```
```{r}
HPG_test <- read_excel("C:\\Users\\ADMIN\\RProjects\\Viet_Nam_Steel_stock_analysis-Time-Series-\\Stock Price - HPG - test.xlsx")
HSG_test <- read_excel("C:\\Users\\ADMIN\\RProjects\\Viet_Nam_Steel_stock_analysis-Time-Series-\\Stock Price - HSG - test.xlsx")
NKG_test <- read_excel("C:\\Users\\ADMIN\\RProjects\\Viet_Nam_Steel_stock_analysis-Time-Series-\\Stock Price - NKG - test.xlsx")
```
```{r}
rmse(m.varf_lev_ft[500:509,1], HPG_test$`<Close>`)
mape(m.varf_lev_ft[500:509,1], HPG_test$`<Close>`) * 100
```
```{r}
rmse(m.varf_lev_ft[500:509,2], HSG_test$`<Close>`)
mape(m.varf_lev_ft[500:509,2], HSG_test$`<Close>`) * 100
```
```{r}
rmse(m.varf_lev_ft[500:509,3], NKG_test$`<Close>`)
mape(m.varf_lev_ft[500:509,3], NKG_test$`<Close>`) * 100
```



Now, click the **Run** button on the chunk toolbar to [execute](https://www.jetbrains.com/help/pycharm/r-markdown.html#run-r-code) the chunk code. The result should be placed under the chunk.
Click the **Knit and Open Document** to build and preview an output.
