---
title: "Time series quarterly analysis"
output: pdf_document
date: "2023-03-23"
---
# import library
```{r setup, include=FALSE}
# truoc khi thay data + run all can sua data dung format, sua start_q la thoi gian bat dau, sua model 3, model 8 neu data khong am
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)
library(tidyverse)
library(Metrics)
library(astsa)
library(tseries)
library(readxl)
library(tsutils)
library(forecast)
```

# import data + plot series:
```{r}
clean_hpg <- read_excel("D:/Time series/clean_hpg_full.xlsx")

start_q <- 2007 #2007Q1


ts_netp <- ts(clean_hpg$`Net Profit`, start = c(start_q, 1), freq = 4)
t <- seq_along(ts_netp)
# prediction data for 2023
t_pred <- seq(length(t) + 2, length(t) * 2 + 1)
seas <- seasdummy(63, 4) # 63 = length(ts)
predict_data <- data.frame(t = t_pred, seas1 = seas[, 1], seas2 = seas[, 2], seas3 = seas[, 3])


head(clean_hpg)
```

```{r}
plot.ts(ts_netp)
```

# Model 1: Linear time trend (series ~ time)
For each model:
1. fit + summarize model \
2. explain coefficients if possible \
3. Graph actual, fitted value, forecast value \
4. Model evaluation: F-test (overall significant), coefficients' significance, R-square \
5. Comments on RMSE, MAPE \
```{r}
model1 <- lm(ts_netp ~ t)
summary(model1)
```
```{r}
plot(ts_netp, xlim = c(start_q - 1, 2024))
fitted1 <- ts(fitted(model1), start = c(2007, 1), freq = 4)
prediction1 <- ts(predict(model1, newdata = predict_data)[1:4], start = c(2023, 1), freq = 4)
lines(fitted1, col = "red")
lines(prediction1, col = "blue")
```
```{r}
# RMSE, MAPE all observation:
rmse(ts_netp, fitted(model1))
mape(ts_netp, fitted(model1))
```
```{r}
AIC(model1)
```

```{r}
# RMSE, MAPE last 4 observations:
rmse(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model1)[(length(ts_netp) - 3): length(ts_netp)])
mape(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model1)[(length(ts_netp) - 3): length(ts_netp)])
```



# Model 2: Log time trend (series ~ log(time))
```{r}

model2 <- lm(ts_netp ~ log(t))
summary(model2)
```
```{r}
plot(ts_netp, xlim = c(start_q - 1, 2024))
fitted2 <- ts(fitted(model2), start = c(2007, 1), freq = 4)
prediction2 <- ts(predict(model2, newdata = predict_data)[1:4], start = c(2023, 1), freq = 4)
lines(fitted2, col = "red")
lines(prediction2, col = "blue")
```
```{r}
# RMSE, MAPE all observation:
rmse(ts_netp, fitted(model2))
mape(ts_netp, fitted(model2))
```
```{r}
AIC(model2)
```

```{r}
# RMSE, MAPE last 4 observations:
rmse(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model2)[(length(ts_netp) - 3): length(ts_netp)])
mape(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model2)[(length(ts_netp) - 3): length(ts_netp)])
```

# Model 3: log-linear time trend (log(series) ~ time)
```{r}
# because net profit series have negative values (indicating loss), cannot take log -> rescale data
net_p_positive <- ts_netp + abs(min(ts_netp)) + 1
model3 <- lm(log(net_p_positive) ~ t)
summary(model3)
```

```{r}
plot(ts_netp, xlim = c(start_q - 1, 2024))
fitted3 <- ts(exp(fitted(model3)) - abs(min(ts_netp)) - 1, start = c(2007, 1), freq = 4)
prediction3 <- ts(exp(predict(model3, newdata = predict_data)[1:4]) - abs(min(ts_netp)) - 1, start = c(2023, 1), freq = 4)
lines(fitted3, col = "red")
lines(prediction3, col = "blue")
```

```{r}
# RMSE, MAPE all observation:
rmse(ts_netp, fitted(model3))
mape(ts_netp, fitted(model3))
```
```{r}
AIC(model3)
```


```{r}
# RMSE, MAPE last 4 observations:
rmse(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model3)[(length(ts_netp) - 3): length(ts_netp)])
mape(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model3)[(length(ts_netp) - 3): length(ts_netp)])
```

# Model 4: seasonal (series ~ s2 + s3 +s4)

```{r}

model4 <- lm(ts_netp ~ seas)
summary(model4)
```

```{r}
plot(ts_netp, xlim = c(start_q - 1, 2024))
fitted4 <- ts(fitted(model4), start = c(2007, 1), freq = 4)
prediction4 <- ts(predict(model4, newdata = predict_data)[1:4], start = c(2023, 1), freq = 4)
lines(fitted4, col = "red")
lines(prediction4, col = "blue")
```

```{r}
# RMSE, MAPE all observation:
rmse(ts_netp, fitted(model4))
mape(ts_netp, fitted(model4))
```

```{r}
AIC(model4)
```


```{r}
# RMSE, MAPE last 4 observations:
rmse(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model4)[(length(ts_netp) - 3): length(ts_netp)])
mape(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model4)[(length(ts_netp) - 3): length(ts_netp)])
```

# Model 5: linear trend + seasonal (series ~ time + seas)
```{r}

model5 <- lm(ts_netp ~ t + seas)
summary(model5)
```
```{r}
plot(ts_netp, xlim = c(start_q - 1, 2024))
fitted5 <- ts(fitted(model5), start = c(2007, 1), freq = 4)
prediction5 <- ts(predict(model5, newdata = predict_data)[1:4], start = c(2023, 1), freq = 4)
lines(fitted5, col = "red")
lines(prediction5, col = "blue")
```

```{r}
# RMSE, MAPE all observation:
rmse(ts_netp, fitted(model5))
mape(ts_netp, fitted(model5))
```

```{r}
AIC(model5)
```

```{r}
# RMSE, MAPE last 4 observations:
rmse(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model5)[(length(ts_netp) - 3): length(ts_netp)])
mape(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model5)[(length(ts_netp) - 3): length(ts_netp)])
```

# Model 6: linear time trend * seasonal (series ~ time * seas)
```{r}

model6 <- lm(ts_netp ~ t * seas)
summary(model6)
```

```{r}
plot(ts_netp, xlim = c(start_q - 1, 2024))
fitted6 <- ts(fitted(model6), start = c(2007, 1), freq = 4)
prediction6 <- ts(predict(model6, newdata = predict_data)[1:4], start = c(2023, 1), freq = 4)
lines(fitted6, col = "red")
lines(prediction6, col = "blue")
```

```{r}
# RMSE, MAPE all observation:
rmse(ts_netp, fitted(model6))
mape(ts_netp, fitted(model6))
```
```{r}
AIC(model6)
```

```{r}
# RMSE, MAPE last 4 observations:
rmse(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model6)[(length(ts_netp) - 3): length(ts_netp)])
mape(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model6)[(length(ts_netp) - 3): length(ts_netp)])
```

# Model 7: Holt-winters additive
```{r}

model7 <- hw(ts_netp, seasonal = "a") # package forecast
summary(model7)
```
```{r}
prediction7 <- predict(model7)$mean
plot(ts_netp, xlim = c(start_q - 1, 2024))
lines(fitted(model7), col = "red")
lines(ts(prediction7[2:5], start = c(2023, 1), freq = 4), col = "blue")
```

```{r}
# RMSE, MAPE all observation:
rmse(ts_netp, fitted(model7))
mape(ts_netp, fitted(model7))
```


```{r}
# RMSE, MAPE last 4 observations:
rmse(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model7)[(length(ts_netp) - 3): length(ts_netp)])
mape(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model7)[(length(ts_netp) - 3): length(ts_netp)])
```
# Model 8: Holt-winters multiplicative
```{r}

model8 <- HoltWinters(ts_netp, seasonal = "m")
model8
```
```{r}
prediction8 <- predict(model8, 10)
plot(ts_netp, xlim = c(start_q - 1, 2024))
lines(fitted(model8)[, 1], col = "red")
lines(ts(prediction8[2:10], start = c(2023, 1), freq = 4), col = "blue")
```

```{r}
# RMSE, MAPE all observation:
rmse(ts_netp, fitted(model8))
mape(ts_netp, fitted(model8))
```


```{r}
# RMSE, MAPE last 4 observations:
rmse(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model8)[(length(ts_netp) - 3): length(ts_netp)])
mape(ts_netp[(length(ts_netp) - 3): length(ts_netp)], fitted(model8)[(length(ts_netp) - 3): length(ts_netp)])
```
# Model 9: decomposition (hard):
# Model selection 
-> Minimize AIC, SIC, HQ \
-> maximize log-likelihood \
-> compare forecast value: \texttt{$Y^F_t = \hat Y _t $} \
-> minimize forecast error: \texttt{$Y^F_t - \hat Y _t $} (MAPE, RMSE)