---
title: "Stock price forecast group"
output: html_document
date: "2023-03-24"
---

```{r setup, include=FALSE}
library(Hmisc)
library(tidyverse)
library(Metrics)
library(readr)
library(astsa)
library(lubridate)
library(tseries)
library(forecast)
library(urca)
library(readxl)

source_functions <- function(fcn_names, file_name) {
    env <- new.env()
    source(file_name, local = env)
    for ( i in 1:length(fcn_names) ) {
        assign(fcn_names[i], get(fcn_names[i], envir = env), envir = .GlobalEnv)
    }
}

source_functions("interp_urdf", "Interprete urdf.R")
ls()

plot_arima <- function(model) {
  plot(forecast(model, h = 10))
  lines(fitted(model), col = "red")
  lines(ts_test, col = "green")
  lines(forecast(model, h = 10)$mean)
}

evaluate_arima <- function(model) {
  autoplot(model)
  checkresiduals(model)
}


rmse_mape_all <- function(model) {
  print("RMSE MAPE on all observations")
  print(rmse(ts_closing, fitted(model)))
  print(mape(ts_closing, fitted(model)))

}
  
rmse_mape_last_10 <- function(model) {
  
  print("RMSE MAPE on last observations")
  print(rmse(ts_closing[(length(ts_closing) - 9): length(ts_closing)], fitted(model)[(length(ts_closing) - 9): length(ts_closing)]))
  print(mape(ts_closing[(length(ts_closing) - 9): length(ts_closing)], fitted(model)[(length(ts_closing) - 9): length(ts_closing)]))

}

rmse_mape_test_10 <-function(model) {
  print("RMSE MAPE on forecasted 10 first days of 2023")
  prediction<- forecast(model, h = 10)$mean
  print(rmse(ts_test[1:10], prediction))
  print(mape(ts_test[1:10], prediction))
}



```
# Import data
```{r}
train <- read_excel("C:\\Users\\ADMIN\\RProjects\\Viet_Nam_Steel_stock_analysis-Time-Series-\\Stock Price - HSG - train.xlsx", col_names = FALSE) #use 2 years 2021 and 2022 to train model
hpg_2023 <- read_excel("C:\\Users\\ADMIN\\RProjects\\Viet_Nam_Steel_stock_analysis-Time-Series-\\Stock Price - HSG - test.xlsx", col_names = FALSE)

ts_closing <- ts(train, start = 2021, frequency = 250)
ts_test <- ts(hpg_2023, start = 2023, frequency = 250)
```

# Step 1: Check for stationarity
## 1.1 Graph the series to quickly detect non-stationarity
```{r}
plot(ts_closing)
lines(ts_test, col = "green")
```

## 1.2 Augmented Dickey-fuller test:

Test with trend and constant (H1: stationary around trend) \
if trend is insignificant: Test with constant (H1: stationary around constant)\
if constant is insignificant: Test without constant (H1: stationary around 0)\

Note: \ 
if trend significant -> detrend series (by diff or others) \
if constant significant -> set include.constant in Arima (package forecast) \
```{r}
test_trend <- ur.df(ts_closing, type = "trend", lag = 0)
summary(test_trend)
interp_urdf(test_trend)
```

```{r}
test_drift <- ur.df(diff(ts_closing), type = "drift", lag = 0) 
summary(test_drift)
interp_urdf(test_drift)
```
```{r}
test_stationarity <- ur.df(diff(ts_closing), type = "none", lag = 0) 
summary(test_stationarity)
interp_urdf(test_stationarity)
```

-> with 1 diff, The series is stationary around constant -> d = 1

# Step 2: determine order for AR and MA (p, q) by checking acf and pacf
Guide: \
Auto regressive (AR) process (p): Order p is the lag value after which PACF plot crosses the upper confidence interval for the first time. \
Moving average (MA) process(q): Order q of the MA process is obtained from the ACF plot, this is the lag after which ACF crosses the upper confidence interval for the first time.\
```{r}
acf(diff(ts_closing))
pacf(diff(ts_closing))
```
auto arima to search for optimize number for p d q
```{r}
fit <- auto.arima(ts_closing,max.p = 5,max.q = 5,max.P = 5,max.Q = 5,max.d = 5,seasonal = TRUE,ic = 'aic')
fit
```
First guess ARIMA(5, 2, 3)
```{r}
test_2_diff <- ur.df(diff(diff(ts_closing)), type = "drift", lag = 0)
summary(test_2_diff)
interp_urdf(test_2_diff)
```
```{r}
test_2_diff <- ur.df(diff(diff(ts_closing)), type = "none", lag = 0)
summary(test_2_diff)
interp_urdf(test_2_diff)
```

```{r}
acf(diff(diff(ts_closing)))
pacf(diff(diff(ts_closing)))
```
# Step 3: fit model
```{r}
arima523 <- Arima(ts_closing, order = c(5, 2, 3), include.constant = TRUE)
summary(arima523 )
```

# Step 4: Evaluate model
## Inversed roots are inside the unit circle
## insignificant coefficient -> remove
## Serial correlation -> add more AR (p)
## Not white noise residuals -> add more MA (q)
```{r}
evaluate_arima(arima523 )
```
```{r}
plot_arima(arima523)
```
```{r}
rmse_mape_all(arima523)
rmse_mape_last_10(arima523)
rmse_mape_test_10(arima523)
```

Good model: Great R-square, log likelihood.\
Small AIC, BIC, HQ forecast error (RMSE, MAPE)\


```{r}
autoplot(arima523)
checkresiduals(arima523)

```





