---
title: "Time series quarterly analysis"
output: pdf_document
date: "2023-03-23"
---
# import library
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)
library(tidyverse)
library(Metrics)
library(astsa)
library(tseries)
library(readxl)
library(tsutils)
library(forecast)
```

# import data + plot series:
```{r}
# enter code here + comments about data
```

# Model 1: Linear time trend (series ~ time)
For each model:
1. fit + summarize model
2. explain coefficients if possible
3. Graph actual, fitted value, forecast value
4. Model evaluation: F-test (overall significant), coefficients' significance, R-square
5. Comments on RMSE, MAPE
```{r}
# fit model here
```
# Model 2: Log time trend (series ~ log(time))
```{r}

```
# Model 3: log-linear time trend (log(series) ~ time)
```{r}

```

# Model 4: seasonal (series ~ s2 + s3 +s4)
```{r}
# enter code here
```
# Model 5: linear trend + seasonal (series ~ time + seas)
```{r}
# enter code here
```

# Model 6: linear time trend * seasonal (series ~ time * seas)
```{r}
# enter code here
```

# Model 7: Holt-winters additive
```{r}

```
# Model 8: Holt-winters multiplicative
```{r}

```
# Model 9: decomposition (hard):
# Model selection 

-> Minimize AIC, SIC, HQ \
-> maximize log-likelihood \
-> compare forecast value: \texttt{$Y^F_t = \hat Y_t $} \
-> minimize forecast error: \texttt{$Y^F_t - \hat Y_t $} (MAPE, RMSE)