---
title: "Stock price forecast group"
output: html_document
date: "2023-03-24"
---

```{r setup, include=FALSE}
library(Hmisc)
library(tidyverse)
library(Metrics)
library(readr)
library(astsa)
library(lubridate)
library(tseries)
library(forecast)
```
# Import data
```{r}

```

# Step 1: Check for stationarity
## 1.1 Graph the series to quickly detect non-stationarity
```{r}

```
## 1.2 Augmented Dickey-fuller test:

Test with trend and constant (H1: stationary around trend) \
if trend is insignificant: Test with constant (H1: stationary around constant)\
if constant is insignificant: Test without constant (H1: stationary around 0)\

Note: \ 
if trend significant -> detrend series (by diff or others) \
if constant significant -> set include.constant in Arima (package forecast) \
```{r}

```

-> diff until stationary -> d is the number of times diff

# Step 2: determine order for AR and MA (p, q) by checking acf and pacf
Guide: \
Auto regressive (AR) process (p): Order p is the lag value after which PACF plot crosses the upper confidence interval for the first time. \
Moving average (MA) process(q): Order q of the MA process is obtained from the ACF plot, this is the lag after which ACF crosses the upper confidence interval for the first time.\
```{r}

```
# Step 3: fit model
```{r}

```

# Step 4: Evaluate model
## Inversed roots are inside the unit circle
## insignificant coefficient -> remove
## Serial correlation -> add more AR (p)
## Not white noise residuals -> add more MA (q)
```{r}

```

Good model: Great R-square, log likelihood.\
Small AIC, BIC, HQ forecast error (RMSE, MAPE)\
# if model not good -> repeat step 2

