---
title: "Time series quarterly analysis"
output: pdf_document
date: "2023-03-23"
---
# import library
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)
library(tidyverse)
library(Metrics)
library(astsa)
library(tseries)
library(readxl)
library(tsutils)
library(forecast)
```
# import data + plot series:
```{r}
# Load Data
df <- read_excel("C:\\Users\\ADMIN\\RProjects\\Viet_Nam_Steel_stock_analysis-Time-Series-\\Gross Sale Revenues - HSG.xlsx")
df
```
```{r}
graph_predict <- function( model, df,start_year, df_predict, start_index, end_index, predict_year){
    plot(df, xlim = c(start_year - 1, 2024))
    fitted <- ts(model$fitted, start = c(start_year, 1), freq = 4)
    prediction <- ts(predict(model, newdata = df_predict)[start_index : end_index], start = c(predict_year, 1), freq = 4)
    lines(fitted, col = "red")
    lines(prediction, col = "blue")
}

rmse_mape <- function( model, df){
    rmse <- rmse(df, model$fitted)
    mape <- mape(df, model$fitted) * 100
    result <- data.frame(rmse, mape)
    return(result)
}

rmse_mape_4_last <- function( model, df){
    rmse <- rmse(df[(length(df) - 3): length(df)], model$fitted[(length(df) - 3): length(df)])
    mape <- mape(df[(length(df) - 3): length(df)], model$fitted[(length(df) - 3): length(df)]) * 100
    result <- data.frame(rmse, mape)
    return(result)
}
```
```{r}
gross_sale_revenue <- ts(df$Gross_Sale_Revenues, start = c(2008,1), frequency = 4)
time <- seq(gross_sale_revenue)
summary(time)
```
```{r}
# create dummy season
s1 <- rep(c(1,0,0,0), 14)
s1 <- c(s1, c(1,0,0))

s2 <- rep(c(0,1,0,0), 14)
s2 <- c(s2, c(0,1,0))

s3 <- rep(c(0,0,1,0), 14)
s3 <- c(s3, c(0,0,1))

s4 <- rep(c(0,0,0,1), 14)
s4 <- c(s4, c(0,0,0))
```
```{r}
time_predict <- seq(60, 118)
data_predict <- data.frame(time = time_predict, seas2 = s3 , seas3= s4, seas4= s1)
data_predict
```

```{r}
plot.ts(gross_sale_revenue)
```

# Model 1: Linear time trend (series ~ time)
For each model:
1. fit + summarize model \
2. explain coefficients if possible \
3. Graph actual, fitted value, forecast value \
4. Model evaluation: F-test (overall significant), coefficients' significance, R-square \
5. Comments on RMSE, MAPE \
```{r}
model1 <- lm(gross_sale_revenue ~ time)
summary(model1)
```
```{r}
graph_predict( model1, gross_sale_revenue, 2008, data_predict, 2, 5, 2023)
```

```{r}
rmse_mape(model1, gross_sale_revenue)
```
```{r}
rmse_mape_4_last(model1, gross_sale_revenue)
```
# Model 3: log-linear time trend (log(series) ~ time)
```{r}
model3 <- lm(log(gross_sale_revenue) ~ time)
summary(model3)
```

```{r}
plot(gross_sale_revenue, xlim = c(2008 - 1, 2024), ylim= c(0, 20000000), main="Model 2: log-linear time trend  ")
fitted <- ts(exp(model3$fitted), start = c(2008, 1), freq = 4)
prediction <- ts(exp(predict(model3, newdata = data_predict)[2:5]), start = c(2023, 1), freq = 4)
lines(fitted, col = "red")
lines(prediction, col = "blue")
legend("topleft", legend=c("Actual", "Fitted", "Forecast 2023"),
       col=c("black", "red", "blue"), lty=1:2, cex=0.8)
```
```{r}
rmse(gross_sale_revenue, exp(model3$fitted))
mape(gross_sale_revenue, exp(model3$fitted)) * 100
```
```{r}
rmse(gross_sale_revenue[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)], exp(model3$fitted[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)]))
mape(gross_sale_revenue[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)], exp(model3$fitted[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)])) * 100
```

# Model 4: seasonal (series ~ s2 + s3 +s4)

```{r}

model4 <- lm(gross_sale_revenue ~ time + s2 +s3 + s4)
summary(model4)
```

```{r}
graph_predict(model4, gross_sale_revenue, 2008, data_predict, 2, 5, 2023)
```

```{r}
# RMSE, MAPE all observation:
rmse_mape(model4, gross_sale_revenue)
```
```{r}
# RMSE, MAPE last 4 observations:
rmse_mape_4_last(model4, gross_sale_revenue)
```
# Model 5: linear trend + seasonal (series ~ time + seas)
```{r}
model5 <- lm(log(gross_sale_revenue) ~ time + s2 +s3 + s4)
summary(model5)
```
```{r}
plot(gross_sale_revenue, xlim = c(2008 - 1, 2030), ylim= c(0, 20000000), main="Model 4: log linear trend + seasonality ")
fitted <- ts(exp(model5$fitted), start = c(2008, 1), freq = 4)
prediction <- ts(exp(predict(model5, newdata = data_predict)[2:5]), start = c(2023, 1), freq = 4)
lines(fitted, col = "red")
lines(prediction, col = "blue")
legend("topleft", legend=c("Actual", "Fitted", "Forecast 2023"),
       col=c("black", "red", "blue"), lty=1:2, cex=0.8)
```
```{r}
rmse(gross_sale_revenue, exp(model5$fitted))
mape(gross_sale_revenue, exp(model5$fitted)) * 100
```
```{r}
rmse(gross_sale_revenue[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)], exp(model5$fitted[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)]))
mape(gross_sale_revenue[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)], exp(model5$fitted[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)])) * 100
```

# Model 6: linear time trend * seasonal (series ~ time * seas)
```{r}
model6 <- lm(gross_sale_revenue ~ time + time * s2 + time * s3 + time * s4)
summary(model6)
```

```{r}
graph_predict(model6, gross_sale_revenue, 2008, data_predict, 2, 5, 2023)
```
```{r}
# RMSE, MAPE all observation:
rmse_mape(model6, gross_sale_revenue)
```
```{r}
# RMSE, MAPE last 4 observations:
rmse_mape_4_last(model6, gross_sale_revenue)
```
# Model 7: Holt-winters additive
```{r}
model7 <- hw(gross_sale_revenue, seasonal = "a") # package forecast
summary(model7)
```
```{r}
prediction7 <- predict(model7)$mean
plot(gross_sale_revenue, xlim = c(2008 - 1, 2024))
lines(fitted(model7), col = "red")
lines(ts(prediction7[2:5], start = c(2023, 1), freq = 4), col = "blue")
```

```{r}
# RMSE, MAPE all observation:
rmse_mape(model7, gross_sale_revenue)
```
```{r}
# RMSE, MAPE last 4 observations:
rmse_mape_4_last(model7, gross_sale_revenue)
```
# Model 8: Holt-winters multiplicative
```{r}
model8 <- hw(gross_sale_revenue, seasonal = "m")
model8
```
```{r}
prediction8 <- predict(model8)$mean
plot(gross_sale_revenue, xlim = c(2008 - 1, 2024))
lines(fitted(model7), col = "red")
lines(ts(prediction7[2:5], start = c(2023, 1), freq = 4), col = "blue")
```

```{r}
# RMSE, MAPE all observation:
rmse_mape(model8, gross_sale_revenue)
```
```{r}
# RMSE, MAPE last 4 observations:
rmse_mape_4_last(model8, gross_sale_revenue)
```
# Model 9: decomposition (hard):
# Model selection
# try to decompose trend, seasonality, and random then do regression with trend or seasonality or trend + seasonality
```{r}
decom.gross_sale_revenue.a <- decompose(gross_sale_revenue, type = 'additive')
decom.gross_sale_revenue.a.trend <- decom.gross_sale_revenue.a$trend
decom.gross_sale_revenue.a.seas <- decom.gross_sale_revenue.a$seasonal
decom.gross_sale_revenue.a.rand <- decom.gross_sale_revenue.a$rand
decom.gross_sale_revenue.a.trendseas <- gross_sale_revenue - decom.gross_sale_revenue.a$rand
decom.gross_sale_revenue.a.seasadj <- gross_sale_revenue - decom.gross_sale_revenue.a$seasonal
plot(decom.gross_sale_revenue.a)
```
```{r}
decom.gross_sale_revenue.a$trend
```
```{r}
decom.gross_sale_revenue.a$seasonal
```
```{r}
plot(gross_sale_revenue, col = 'blue')
lines(decom.gross_sale_revenue.a.trend)
lines(decom.gross_sale_revenue.a.trendseas, col = 'red')
```

```{r}
decom.gross_sale_revenue.a.seasadj <- gross_sale_revenue - decom.gross_sale_revenue.a$seasonal
model9.seasadj = lm(decom.gross_sale_revenue.a.seasadj ~ time )
summary(model9.seasadj)
```
```{r}
predict(model9.seasadj, newdata = data_predict)[2 : 5] + decom.gross_sale_revenue.a.seas[1:4]
pred <- predict(model9.seasadj, newdata = data_predict)[2 : 5] + decom.gross_sale_revenue.a.seas[1:4]
prediction <- ts(pred, start = c(2023, 1), freq = 4)
prediction
```
```{r}
plot(gross_sale_revenue, xlim = c(2008 - 1, 2024), main="Model 8: Additive seasonal adjustment ~ time")
fitted <- ts(model9.seasadj$fitted + decom.gross_sale_revenue.a.seas, start = c(2008, 1), freq = 4)
lines(fitted, col = "red")
lines(prediction, col = "blue")
legend("topleft", legend=c("Actual", "Fitted", "Forecast 2023"),
       col=c("black", "red", "blue"), lty=1:2, cex=0.8)
```
```{r}
rmse(gross_sale_revenue, model9.seasadj$fitted + decom.gross_sale_revenue.a.seas)
mape(gross_sale_revenue, model9.seasadj$fitted + decom.gross_sale_revenue.a.seas) * 100
```
```{r}
rmse(gross_sale_revenue[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)], (model9.seasadj$fitted + decom.gross_sale_revenue.a.seas)[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)])
mape(gross_sale_revenue[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)], (model9.seasadj$fitted + decom.gross_sale_revenue.a.seas)[(length(gross_sale_revenue) - 3): length(gross_sale_revenue)]) * 100
```



